---
title: "Tooth Growth Analysis"
author: "Ran Ju"
output:
  pdf_document:
    toc: false
---
# Exploratory Data Analysis
```{r}
# Load library and data
library(ggplot2)
data(ToothGrowth)
tg <- ToothGrowth
# Fix the seed for reproducibility
set.seed(344344)
```
```{r}
# Box plot of len against supp
ggplot(data = tg, aes(x = supp, y = len)) +
  geom_boxplot(aes(fill = supp)) +
  geom_jitter() +
  ggtitle("Tooth Length by Supplement Type")
```
```{r}
# Box plot of len against dose
ggplot(data = tg, aes(x = factor(dose), y = len)) +
  geom_boxplot(aes(fill = factor(dose))) +
  geom_jitter() +
  ggtitle("Tooth Length by Dose")
```
```{r}
qplot(dose, len, data=tg, geom=c("point", "smooth"), 
      method="lm", formula=y~x, color=supp,
      main="Regression of Tooth Lengh on Dose", 
      xlab="Dose", ylab="Tooth Length")
```

It seems that orange juice is better for tooth length than ascorbic acid, and the more vitamin C the higher the tooth length.

# Is tooth length related to supplement?
Define $H_0$ as $E[len|supp = OJ] = E[len|supp = VC] = E[len]$ with $\alpha = 0.05$.  Use resampling to create 10000 simulations, and compute the difference of mean of OJ and VC for each sample.
```{r}
n = dim(tg)[1]
B = 10000
means <- NULL
for (i in 1:B) {
  # Sample with replacement
  ri <- sample.int(n, n, replace = T)
  t <- tg[ri, ]
  # Compute the difference of the mean values
  means <- c(means, mean(t[t[, "supp"] == "VC", "len"]) - mean(t[t[, "supp"] == "OJ", "len"]))
}
ggplot() +
  geom_histogram(aes(x = means), alpha = 0.5, binwidth = 0.25) +
  geom_vline(aes(xintercept = median(means)), color = "red") +
  ggtitle("Distribution of mean(len of VC) - mean(len of OJ)")
# Compute the median
median(means)
```
The plot shows the distribution of the means (the vertical line is the median), and it is less than 0 with probability
```{r}
# Compute the weight of mean >= 0
sum(means >= 0) / B
```
Because the probability is less than $\alpha = 0.05$, we reject $H_0$.

# Is tooth growth related to dose?
We perform a t-test between each pair of dose levels.  
```{r}
t05 <- tg[tg$dose == 0.5, "len"]
t1 <- tg[tg$dose == 1, "len"]
t2 <- tg[tg$dose == 2, "len"]
```
```{r}
# Compare dose = 0.5 and 1
t.test(t05, t1, alternative = "less")
```
```{r}
# Compare dose = 1 and 2
t.test(t1, t2, alternative = "less")
```
```{r}
# Compare dose = 0.5 and 2
t.test(t05, t2, alternative = "less")
```
The p-values are all well below $\alpha = 0.05$, hence we conclude that VC dose has a positive effect on tooth growth, assuming _len_ has a normal distribution in each group.

# Conclusion
Tooth length is higher among the individuals fed with orange juice than ascorbic acid, and VC dose has a a positive correlation with tooth growth.  However, more analysis is need to quantify the relationship, which requires regression analysis.